let's load the packages 

```{r}
library(dplyr)
library(janitor)
library(tidyr)
```


first load the relevant datasets

```{r}
load("./RDATA/titanic_test_sub2.RData")
load("./RDATA/titanic_train.RData")

```
the above load was required because it was first time but after i worked a bit on it, instead of opening up the above  i can straight open the file

	```{r lc, echo = FALSE }
	    load("./RDATA/titanic_child.RData")
	```

let's dig into the age variable

```{r}
summary(titanic.train$Age)
```

let's create a variable whether he is child or not

```{r}
titanic.train$child <- 0
titanic.train$child[titanic.train$Age < 18] <- 1
```

let's check how the child created with age NA

```{r}
titanic.train %>% 
  filter(is.na(Age)) %>%
  select(child,Age)
```

so the above makes sense since we will eventually replace age na with median value

what we would want to check is survival between child and non-child

```{r}
prop.table(table(titanic.train$child,titanic.train$Survived,dnn = c("Child","Survive")),1)
```

the above cross table was created using prop table but i did not quite like it cause i don't like the bast R solution. 

let me digress a bit and look into some alternatives

```{r}
titanic.train %>%
  count(child,Survived) %>% 
  group_by(child) %>%
  mutate(prop=n/sum(n)) %>%
  select(-n) %>%
  spread(Survived,prop)
```


now there is a janitor package which is piping allowed and awesome

```{r}
#require(janitor)
titanic.train %>%
  crosstab(child, Survived) %>%
  adorn_crosstab(digits=1,show_totals = F,show_n=F)
```
	


but the strange thing above is that being child survival prob is not that high but it was really high when female, may be it will be for much younger 

i forgot let me quickly check whether that was true

```{r , echo = FALSE }
titanic.train %>%
    crosstab(Sex, Survived) %>% 
    adorn_crosstab(digits=1, show_totals= F, show_n=F)
```
so in the above confirmed that being women it seems quite to survive than being child which is quite strange
but we will check that out later, for now just let go with the things


First let me replace missing age with median age

at this point let first save the titatnic train file with all these changes

```{r , echo = FALSE }

titanic.train.sub3  <-   titanic.train
save(titanic.train.sub3, file="./RDATA/ttrain_sub3.RData")
```
now time to do the actual replace operation

so there are actually `{r sum(is.na(titanic.train.sub3$Age))}` number of missing values


here we go

```{r}
titanic.train.sub3 <- titanic.train.sub3 %>%
  mutate(Age=replace(Age,is.na(Age),median(Age,na.rm=T)))
```

but it seems that there might be a problem cause there must be earlier defined some child who have been assigned higher age
let's check whether we get something like that or not

```{r}
titanic.train.sub3 %>%
  filter(child==1,Age> 18)
  
```

its not the case because none of the child had missing age



Now what i do is that create a age wise category of survival

```{r}
titanic.child <- titanic.train.sub3 %>%
  mutate(age_cat=cut(Age,
                     breaks=c(seq(0,60,by=10),90),
                     labels=paste(c(paste(seq(1,60,by=10),
                                  seq(10,60,by=10),
                                  sep="-"),"Above 60"),"years")))


save(titanic.child,file="./RDATA/titanic_child.RData")
```

Now let's try to find an idea about the distribution of the age category


```{r ac, echo = FALSE }
    titanic.child %>% 
	tabyl(age_cat)
```			


Now let's find the category of age and survival rate


```{r}
age.survive <- prop.table(table(titanic.child$age_cat,titanic.child$Survived),1)
```

can prop.table be transformed into matrix? turns out yes it is

```{r}
age.survive.mat <- as.matrix(age.survive)
```

yes it does, then can it go to df?

```{r}
as.data.frame(as.matrix(age.survive))
```

following is the way how to transform

```{r}
age.survive.mat <- as.matrix(age.survive)
age.survive.df <- data.frame(NS=age.survive.mat[,1],S=age.survive.mat[,2])
save(age.survive.df,file="./RDATA/age_survive_df.RData")
```


how to do the above with crosstab, let me try in the following 

```{r , echo = FALSE }
age.surv <- titanic.child %>%
    crosstab(age_cat,Survived) %>%
    adorn_crosstab(digits=1,show_totals=F,show_n=F)
```


Let's rename the Survived factor instead of 0 and 1

```{r res, echo = FALSE }
titanic.child =titanic.child %>%
    mutate(Survived=factor(Survived))
levels(titanic.child$Survived)=c("NS","S")
```
save titanic child once again

```{r s2, echo = FALSE }
save(titanic.child, file="./RDATA/titanic_child.RData")
```
now ggplot time

```{r}
#load("./RDATA/age_survive_df.RData")
ggplot(age.surv,aes(x=age_cat,y=S))+
  geom_bar(stat = "identity",fill="blue")

```


with additional variables, the titanic data set is saved in titanic_add.RData. So from now we can directly load it. 

```{r}
#load("./RDATA/titanic_add.RData")
#load("./RDATA/age_survive_df.RData")
```

in the above even the titanic_add is used i would just use the titanic.child since in this new 
round i have been using it

so i will do the following

```{r , echo = FALSE }
titanic.add <- titanic.child
```

now i would like to see those who survived among them what is the age distribution

```{r}
ggplot(titanic.add,aes(x=age_cat,fill=as.character(Survived)))+geom_bar()

```

But the above correctly does not portray, b/c the child seems to be lowest but they may be lowest to start with, how aobut taking proportion

```{r}
ggplot(titanic.add,aes(x=age_cat,fill=as.character(Survived)))+
  geom_bar(position="fill")
```

Now from the above it is quite obvious that younger age category has higher rate of survival, there is more than 50 percent 
chance that some one survive if he/she belonged to 1-10 age category


among the persons survived what is the male female ratio in different age-category?


```{r}
ggplot(filter(titanic.add,
              Survived==1),aes(x=age_cat,fill=Sex))+
  geom_bar(position="fill")
```

the above does not give us much of an information, in each categor female had higher chance of survival and being child female
did not give any added advantage. 


now moving into a little bit of off-topic, i just want to get hold of calculating proportion with dplyr. on the second thought, usual way doing n/sum sounds fine, there is that count function which is useful





now let's look into other variables, there's one is SibSp number of sibilings present, now what i want to do quickly is 
check if there was no siblings does the make prediction better or not

```{r}
titanic.add$sibling <- cut(titanic.add$SibSp,breaks = c(-1,0,8),labels = c("No siblings", "Has siblings"))
```

now check the proportion survived

```{r}
aggregate(Survived ~ sibling,data=titanic.add,mean)
```
the above is not working since all the variables have become  qualitative
we do the same with dplyr version

```{r}
group_by(titanic.add,sibling) %>%
  summarise(mean(Survived))
  
```
 the above is also not working because of categorical variable

its better idea to do this 


```{r , echo = FALSE }
titanic.add %>% 
    crosstab(sibling,Survived) %>%
	adorn_crosstab()
```
in the above we have find that those who have sibling has higher probability of survival


let's have a look into Pclass variable which identifies different classes of tickets

```{r}
table(titanic.add$Pclass)
```

let's check the survival rate among these different classes

```{r}
group_by(titanic.add, Pclass,Survived) %>% 
  summarise(N=n()) %>%
  mutate(perc=N/sum(N))
```

I think a easier version will be

```{r}
group_by(titanic.add, Pclass) %>% 
  summarise(perc=mean(Survived))
```

that's much cleaner with the same results

anyway it seems someone with the third class has pretty low chance of surviving

the above will not work because of categorical variable

so the following will work

```{r , echo = FALSE }
titanic.add %>% 
    crosstab(Pclass,Survived) %>%
    adorn_crosstab()
```
so we see that class three has very low rate of survival

now let's disaggregate this gender wise

```{r}
titanic.add %>%
  count(Pclass,Sex) %>%
  mutate(P=n/sum(n))
```

now the above is simply showing the proportion of male and female but what we want to know is survival in each category

```{r}
titanic.add %>%
group_by( Pclass,Sex) %>% 
  summarise(P=mean(Survived)) 
  
```

now to be explicit i could be doing this

```{r}
titanic.add %>%
count( Pclass,Sex,Survived) %>%
group_by(Pclass,Sex) %>%
mutate(tot=sum(n),
       perc=n/tot) %>%
filter(Survived=="S") %>%
arrange(perc)

```

the above piece of code clearly shows that male or female, Pclass 3 has the lowes chance of survival,
men were the lowest from Pclass 3


in the above we find that the female in the 3rd class had lower chance of survival

now have a look into fare variable, since it's a continuous variable its better to have a summary

```{r}
summary(titanic.add$Fare)
```

let's recode the variable

```{r}
titanic.add$Fare2 <- cut(titanic.add$Fare,
                         breaks = c(-1,10,20,30,515), 
                         labels = c("0-10","11-20","21-30","Above 30"))
```

Now let's have a look at distribution of survival in the fare category

```{r}
titanic.add %>%
    count(Fare2, Survived) %>%
group_by(Fare2) %>%
mutate(tot=sum(n),
       perc=n/tot) %>%
filter(Survived=="S") 
```

In the above we find that lowest fare had the smallest chance of survival

so let's disaggregate across gender

```{r}
titanic.add %>%
    count(Fare2,Sex, Survived) %>%
    group_by(Fare2) %>%
    mutate(tot=sum(n),
	   perc=n/tot) %>%
    filter(Survived=="S") %>%
    arrange(perc)
```

what we find here is that higher fare females has higher chance of survival
than any other category, in general higher fare has higher chance


if we try to visualize then it will be like this

```{r}
ggplot(titanic.add, aes(x=Fare2)) +
  geom_bar(aes(fill=Sex),position="fill")
```

the above is not what i envisioned, the above just gender ratio in each Fare

```{r}
titanic.add %>%
  group_by(Fare2,Sex) %>%
  filter(!is.na(Fare2)) %>%
  summarise(P=mean(Survived)) %>%
  ggplot(.,aes(x=Fare2,y=P)) + geom_bar(stat = "identity",aes(fill=Sex),position = "dodge")
```

in the above we see both male and female has low category of survival

now let's distribute the whole thing in terms of Pclass

```{r}
titanic.add %>%
  group_by(Fare2,Pclass,Sex) %>%
  filter(!is.na(Fare2)) %>%
  summarise(N=n(),P=mean(Survived)) %>%
  arrange(Sex,P)

```

Now time to save the data

```{r}
save(titanic.add, file="./RDATA/titanic_add.RData")
```

Update : 28-06-2017 00:37 ,  so far se have established that Fare and Pclass are important determinant
along with child and being female
titanic.add is the latest file with all the changes required. 




