---
title: "Titanic"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
require(dplyr)
```

```{r}
load("./RDATA/titanic_train.RData")
load("./RDATA/titanic_test.RData")
```
first check survival


```{r}
prop.table(table(titanic$Survived))
```

in the above, we see most of the died, so our best prediction is all died

```{r}
titanic.test$Survived <- 0
submit <- data.frame(PassengerID=titanic.test$PassengerId,Survived=titanic.test$Survived)
write.csv(submit,file="./SUBMIT/theyAllPerish.csv",row.names = F)
```


 another try
let's see whether we can improve

 disaster was famous for saving women and children, let's check what's happening with them


```{r}
prop.table(table(titanic$Sex))
```

what we find is that as 35% was female

let's check how survived across male or female

```{r}
prop.table(table(titanic$Sex,titanic$Survived))
```

but that's not what we wanted, we rather want across row, so we do

```{r}
prop.table(table(titanic$Sex,titanic$Survived),1)
```

that's more like it. 

so we think that most of females survived while males do not, so let's update with female all survived , just a crude prediction

```{r}
titanic.test$Survived[titanic.test$Sex=="female"] <- 1
```

let's just have a quick check whether it worked

```{r}
prop.table(table(titanic.test$Sex,titanic.test$Survived),1)
```

in the above, we see all the female survived and all the male perished. Let's make another submitssion

```{r}
submit <- data.frame(PassengerID=titanic.test$PassengerId,Survived=titanic.test$Survived)
write.csv(submit,file="./SUBMIT/menAllPerish.csv",row.names = F)
```





